---
title: "CDS - The Moods of Migraine"
author: "Cecilie Stilling Pedersen"
date: "1 nov 2021"
output: html_document
---

# Intro: 
What do I want to do?


## Data prep

```{r Packages and WD}
# Load packages
library(dplyr)
library(tidyverse)
library(ggplot2)
library(rvest)
library(tidyr)
library(stringr)
library(janitor)
library(httr)
library(jsonlite)
library(lubridate)

# citation()
# citation("dplyr")
# citation("tidyverse")
# citation("ggplot2")
# Set working directory
setwd("C:/Users/cecil/OneDrive/AU/Cultural Data Science/Exam project")

```

```{r Load data}
# Loading data
mood_data <- read.csv("daylio-export-2021-11-01.csv", header = TRUE, sep = ",")

migraine_data <- read.csv("migraine_data_updated.csv", header = TRUE, sep = ";")

```

### Mood data
We're only interested in the mood data that line up with the migraine data, and we want to convert the registered moods to numbers on a three point scale.
```{r Clean mood_data}
# Check class for full_date
class(mood_data$full_date) # Character

# Change class to date for full_date
mood_data$full_date <- as.Date(mood_data$full_date)

# Check class again to see if it worked
class(mood_data$full_date) # Date

# Remove rows we don't need (we only need the data between 2021-08-20 and 2021-10-31)
mood_data_subset <- mood_data[1:73,]

# Write the weekdays in English
mood_data_clean <- mood_data_subset %>%
  mutate(weekday = if_else(weekday %in% "søndag","sunday",
                   if_else(weekday %in% "lørdag","saturday",
                   if_else(weekday %in% "fredag", "friday",
                   if_else(weekday %in% "torsdag", "thursday",
                   if_else(weekday %in% "onsdag", "wednesday",
                   if_else(weekday %in% "tirsdag", "tuesday",
                   if_else(weekday %in% "mandag", "monday", "other"))))))))

# Get the diffeerent moods in English 
mood_data_clean <- mood_data_clean %>% 
  mutate(mood = if_else(mood %in% "depressed", "depressed",
                if_else(mood %in% "fint", "good",
                if_else(mood %in% "middel", "neutral", 
                if_else(mood %in% "swings", "neutral", "depressed")))))


# Remove the "note" cloumns
mood_data_clean <- mood_data_clean[-c(7,8)]

# Make new column where the different moods are rated on a 1-3 point scale 
mood_data_clean$mood_rating <- mood_data_clean$mood
mood_data_clean <- mood_data_clean %>% 
  mutate(mood_rating = if_else(mood %in% "depressed", "1",
                if_else(mood %in% "good", "3",
                if_else(mood %in% "neutral", "2", 
                if_else(mood %in% "swings", "2", "3")))))


# Check class for mood_rating
class(mood_data_clean$mood_rating) # Character

# Change class to date for full_date
mood_data_clean$mood_rating <- as.numeric(mood_data_clean$mood_rating)

# Check class again to see if it worked
class(mood_data_clean$mood_rating) # numeric


```

### Migraine data

```{r Clean migraine_data}

migraine_data$date <- format(as.Date(migraine_data$date, format = "%d-%m-%Y"), "%Y-%m-%d")

# Check class for date
class(migraine_data$date) # Character

# Change class to date for date
migraine_data$date <- as.Date(migraine_data$date)

# Check class again to see if it worked
class(migraine_data$date) # Date

names(migraine_data)[names(migraine_data) == "date"] <- "full_date"
```

### Merging mood and migraine data
We want to merge the two data sets

```{r Merging mood and migraine data}
# Merging mood_data_clean with migraine_data
data_both <- merge(mood_data_clean, migraine_data)

# create column with migraine, headache and nothing
data_both$pain_type = "nothing"
#data_all$pain_type_value = 0
# 0 = nothing , 1 = headache , 2 = migraine

data_both$pain_type[data_both$migraine > 0] = "migraine"
data_both$pain_type[data_both$headache > 0] = "headache"
# data_all$pain_type_value[data_all$migraine > 0] = 2
# data_all$pain_type_value[data_all$headache > 0] = 1


```

### Weather data
We want to accent weather data for the mood and migraine registering period to see if there is any correlation.

```{r Weather data access}
## IMPORTANT: my weather data license might be outdated, so I have saved the accessed and cleaned data while I had the API key, so it can be loaded anyways - If this is the case, then skip the next two chunks and go directly to the "Merging mood/migraine data and weather data" chunk


# I can only access 35 dates at a time form the worldweatheronline website, so I split it in 3 steps
# 1
weather = GET("http://api.worldweatheronline.com/premium/v1/past-weather.ashx?key=684c5c95ad3844339bf120034211111&q=Aarhus&format=json&date=2021-08-20&enddate=2021-11-15")

weather_text <- content(weather, "text")
weather_text

weather_json <- fromJSON(weather_text, flatten = TRUE)

weather_df <- as.data.frame(weather_json)


# 2
weather2 = GET("http://api.worldweatheronline.com/premium/v1/past-weather.ashx?key=684c5c95ad3844339bf120034211111&q=Aarhus&format=json&date=2021-09-24&enddate=2021-11-15")


weather2_text <- content(weather2, "text")
weather2_text

weather2_json <- fromJSON(weather2_text, flatten = TRUE)

weather2_df <- as.data.frame(weather2_json)


# 3
weather3 = GET("http://api.worldweatheronline.com/premium/v1/past-weather.ashx?key=684c5c95ad3844339bf120034211111&q=Aarhus&format=json&date=2021-10-29&enddate=2021-11-15")

weather3_text <- content(weather3, "text")
weather3_text

weather3_json <- fromJSON(weather3_text, flatten = TRUE)

weather3_df <- as.data.frame(weather3_json)

# Combine the 3 weatherd dataframes using rbind
weather_data <- rbind(weather_df, weather2_df, weather3_df)

# Delete the rows not corresponding to the migraine and mood data
weather_data <- weather_data[-c(74:88), ]

```

```{r Clean weather_data}
# Separeate the lists in the dataframe
weather_data_unlist <- cbind(weather_data[!sapply(weather_data, is.list)], 
                             (t(apply(weather_data[sapply(weather_data, is.list)], 1, unlist))))


### Humidity ###

# Check class
class(weather_data_unlist$data.weather.hourly.humidity1) # character

# Make numeric
weather_data_unlist$data.weather.hourly.humidity1 <- as.numeric(weather_data_unlist$data.weather.hourly.humidity1)

weather_data_unlist$data.weather.hourly.humidity2 <- as.numeric(weather_data_unlist$data.weather.hourly.humidity2)

weather_data_unlist$data.weather.hourly.humidity3 <- as.numeric(weather_data_unlist$data.weather.hourly.humidity3)

weather_data_unlist$data.weather.hourly.humidity4 <- as.numeric(weather_data_unlist$data.weather.hourly.humidity4)

weather_data_unlist$data.weather.hourly.humidity5 <- as.numeric(weather_data_unlist$data.weather.hourly.humidity5)

weather_data_unlist$data.weather.hourly.humidity6 <- as.numeric(weather_data_unlist$data.weather.hourly.humidity6)

weather_data_unlist$data.weather.hourly.humidity7 <- as.numeric(weather_data_unlist$data.weather.hourly.humidity7)

weather_data_unlist$data.weather.hourly.humidity8 <- as.numeric(weather_data_unlist$data.weather.hourly.humidity8)

# Check class again
class(weather_data_unlist$data.weather.hourly.humidity8) # numeric

# Take mean of the hourly humidity rows and put into new column
weather_data_unlist$mean_humidity <- rowMeans(weather_data_unlist[ , 115:122])


### PRESSURE ### 

# Check class
class(weather_data_unlist$data.weather.hourly.pressure1) # character

# Make numeric
weather_data_unlist$data.weather.hourly.pressure1 <- as.numeric(weather_data_unlist$data.weather.hourly.pressure1)

weather_data_unlist$data.weather.hourly.pressure2 <- as.numeric(weather_data_unlist$data.weather.hourly.pressure2)

weather_data_unlist$data.weather.hourly.pressure3 <- as.numeric(weather_data_unlist$data.weather.hourly.pressure3)

weather_data_unlist$data.weather.hourly.pressure4 <- as.numeric(weather_data_unlist$data.weather.hourly.pressure4)

weather_data_unlist$data.weather.hourly.pressure5 <- as.numeric(weather_data_unlist$data.weather.hourly.pressure5)

weather_data_unlist$data.weather.hourly.pressure6 <- as.numeric(weather_data_unlist$data.weather.hourly.pressure6)

weather_data_unlist$data.weather.hourly.pressure7 <- as.numeric(weather_data_unlist$data.weather.hourly.pressure7)

weather_data_unlist$data.weather.hourly.pressure8 <- as.numeric(weather_data_unlist$data.weather.hourly.pressure8)

# Check class again
class(weather_data_unlist$data.weather.hourly.pressure8) # numeric

# Take mean of the hourly humidity rows and put into new column
weather_data_unlist$mean_pressure <- rowMeans(weather_data_unlist[ , 139:146])


### WindGustKmph ###

# Check class
class(weather_data_unlist$data.weather.hourly.WindGustKmph1) # character

# Make numeric
weather_data_unlist$data.weather.hourly.WindGustKmph1 <- as.numeric(weather_data_unlist$data.weather.hourly.WindGustKmph1)

weather_data_unlist$data.weather.hourly.WindGustKmph2 <- as.numeric(weather_data_unlist$data.weather.hourly.WindGustKmph2)

weather_data_unlist$data.weather.hourly.WindGustKmph3 <- as.numeric(weather_data_unlist$data.weather.hourly.WindGustKmph3)

weather_data_unlist$data.weather.hourly.WindGustKmph4 <- as.numeric(weather_data_unlist$data.weather.hourly.WindGustKmph4)

weather_data_unlist$data.weather.hourly.WindGustKmph5 <- as.numeric(weather_data_unlist$data.weather.hourly.WindGustKmph5)

weather_data_unlist$data.weather.hourly.WindGustKmph6 <- as.numeric(weather_data_unlist$data.weather.hourly.WindGustKmph6)

weather_data_unlist$data.weather.hourly.WindGustKmph7 <- as.numeric(weather_data_unlist$data.weather.hourly.WindGustKmph7)

weather_data_unlist$data.weather.hourly.WindGustKmph8 <- as.numeric(weather_data_unlist$data.weather.hourly.WindGustKmph8)

# Check class again
class(weather_data_unlist$data.weather.hourly.WindGustKmph8) # numeric

# Take mean of the hourly humidity rows and put into new column
weather_data_unlist$mean_WindGustKmph <- rowMeans(weather_data_unlist[ , 219:226])


### PRECIPITATION MM ###

# Check class
class(weather_data_unlist$data.weather.hourly.precipMM1) # character

# Make numeric
weather_data_unlist$data.weather.hourly.precipMM1<- as.numeric(weather_data_unlist$data.weather.hourly.precipMM1)

weather_data_unlist$data.weather.hourly.precipMM2 <- as.numeric(weather_data_unlist$data.weather.hourly.precipMM2)

weather_data_unlist$data.weather.hourly.precipMM3 <- as.numeric(weather_data_unlist$data.weather.hourly.precipMM3)

weather_data_unlist$data.weather.hourly.precipMM4 <- as.numeric(weather_data_unlist$data.weather.hourly.precipMM4)

weather_data_unlist$data.weather.hourly.precipMM5 <- as.numeric(weather_data_unlist$data.weather.hourly.precipMM5)

weather_data_unlist$data.weather.hourly.precipMM6 <- as.numeric(weather_data_unlist$data.weather.hourly.precipMM6)

weather_data_unlist$data.weather.hourly.precipMM7 <- as.numeric(weather_data_unlist$data.weather.hourly.precipMM7)

weather_data_unlist$data.weather.hourly.precipMM8 <- as.numeric(weather_data_unlist$data.weather.hourly.precipMM8)

# Check class again
class(weather_data_unlist$data.weather.hourly.precipMM8) # numeric

# Take mean of the hourly humidity rows and put into new column
weather_data_unlist$mean_precipMM <- rowMeans(weather_data_unlist[ , 99:106])


# Save the weather data so we have access to it even after the license is outdated
write.csv(weather_data_unlist,"weather_data_unlist.csv", row.names = FALSE)



# Create dataframe with the columns I want to use

weather_data_select <- weather_data_unlist %>% 
  select("data.weather.date", "data.weather.maxtempC", "data.weather.mintempC", "data.weather.avgtempC", "data.weather.sunHour", "mean_humidity", "mean_pressure", "mean_WindGustKmph", "mean_precipMM")

weather_data_select <- weather_data_select %>% 
  rename(full_date = data.weather.date)


# Save the weather data so we have access to it even after the license is outdated
# write.csv(weather_data_select,"weather_data_select.csv", row.names = FALSE)

```


### Merging the data
We want to merge the weather data with the mood/migriane data

```{r Merging mood/migraine data and weather data}
# In case the API keys has been diabled, load the weater data 
weather_data_select <- read.csv("weather_data_select.csv", header = TRUE, sep = ",")

# Merging mood_data_clean with migraine_data
data_all <- merge(data_both, weather_data_select, by = "full_date")

# Merging does not work - it returns the columns but 0 rows, so we inspect the column to be joined by
"full_date" %in% names(data_both) #TRUE
"full_date" %in% names(weather_data_select) #TRUE

# Check class for full_date in weather data
class(weather_data_select$full_date) #character

# The issue is that ful_date in the weather data is not dates but character
# Change class to date for full_date
weather_data_select$full_date <- as.Date(weather_data_select$full_date)

# Check class again to see if it worked
class(weather_data_select$full_date) # Date

# Merging mood_data_clean with migraine_data
data_all <- merge(data_both, weather_data_select, by = "full_date")


# Save the final data set
# write.csv(data_all,"data_all.csv", row.names = FALSE)

```



## Data visualization

```{r Distribution overview}

# Bar plot to give an overview of the general mood of Patient X
ggplot(data_all, aes(x = mood, fill = mood))+
  geom_bar() 




# Bar plot to give an overview of the  migraine and headaches
ggplot(data_all, aes(x=pain_type, fill=pain_type)) +
  geom_bar()


```

```{r Visualizations}

# Plot to see the ratings of mood and migraine throughout the registered period
ggplot(data_all, aes(full_date)) + 
  geom_line(aes(y = migraine, colour = "migraine")) + 
  geom_line(aes(y = mood_rating, colour = "mood_rating"))

  # It looks very much as expected. The two loines are mostly opposite of each other. When there is a        migraine, the mood is down and vice versa. 



# Plotting an overview of the migraines throughout the registered period. A smoothing line is added in order to see what the trends look like. Note that the geom_smooth function adds confidence bands on the smooth as well. 
ggplot(data_all, aes(full_date)) + 
  geom_smooth(aes(y = migraine, colour = "migraine"))

# Plotting an overview of the regular headaches throughout the registered period. A smoothing line is added in order to see what the trends look like. Note that the geom_smooth function adds confidence bands on the smooth as well. 
ggplot(data_all, aes(full_date)) +   
  geom_smooth(aes(y = headache, colour = "headache"))

# A comparison of the smoothed trends, just for funsies. 
ggplot(data_all, aes(full_date)) +   
  geom_smooth(aes(y = headache, colour = "headache"))+
  geom_smooth(aes(y = migraine, colour = "migraine"))




# Plotting an overview of the mood rating throughout the registered period. A smoothing line is added in order to see what the trends look like. Note that the geom_smooth function adds confidence bands on the smooth as well. 
ggplot(data_all, aes(full_date)) + 
  geom_smooth(aes(y = mood_rating, colour = "mood_rating"))


# Mood, migraines, and headaches plotted together for a comparison
ggplot(data_all, aes(full_date)) + 
  geom_smooth(aes(y = mood_rating, colour = "mood_rating"))+
  geom_smooth(aes(y = headache, colour = "headache"))+
  geom_smooth(aes(y = migraine, colour = "migraine"))
  
  

# Plotting an overview of the humidity throughout the registered period. A smoothing line is added in order to see what the trends look like. Note that the geom_smooth function adds confidence bands on the smooth as well.
ggplot(data_all, aes(full_date)) + 
  geom_smooth(aes(y = mean_humidity, colour = "mean_humidity"))

  
# Looking at the mean humidity together with mood and migraine (scaled)
scaleFactor <- max(data_all$mood_rating) / max(data_all$mean_humidity)

ggplot(data_all, aes(x=full_date)) +
  geom_smooth(aes(y=migraine), method="loess", col="blue") +
  geom_smooth(aes(y=mood_rating), method = "loess", col="green") +
  geom_smooth(aes(y=mean_humidity * scaleFactor), method="loess", col="red") +
  scale_y_continuous(name="Migraine and mood", sec.axis=sec_axis(~./scaleFactor, name="Humidity")) +
  theme(
    axis.title.y.left=element_text(color="blue"),
    axis.text.y.left=element_text(color="blue"),
    axis.title.y.right=element_text(color="red"),
    axis.text.y.right=element_text(color="red")
  )

# Plotting an overview of the avg temp throughout the registered period. A smoothing line is added in order to see what the trends look like. Note that the geom_smooth function adds confidence bands on the smooth as well.
ggplot(data_all, aes(full_date)) + 
  geom_smooth(aes(y = data.weather.avgtempC, colour = "data.weather.avgtempC"))

# Looking at the avg temp together with mood and migraine (scaled)
scaleFactor1 <- max(data_all$mood_rating) / max(data_all$data.weather.avgtempC)

ggplot(data_all, aes(x=full_date)) +
  geom_smooth(aes(y=migraine), method="loess", col="blue") +
  geom_smooth(aes(y=mood_rating), method = "loess", col="green") +
  geom_smooth(aes(y=data.weather.avgtempC * scaleFactor1), method="loess", col="red") +
  scale_y_continuous(name="Migraine and mood", sec.axis=sec_axis(~./scaleFactor1, name="Avg Temp")) +
  theme(
    axis.title.y.left=element_text(color="blue"),
    axis.text.y.left=element_text(color="blue"),
    axis.title.y.right=element_text(color="red"),
    axis.text.y.right=element_text(color="red")
  )

```



